---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
all_deaths
```

```{r}
nrow(all_deaths)
```
```{r}
ncol(all_deaths)
```

```{r}
dim(all_deaths)
```

```{r}
names(all_deaths)
```

```{r}
head(all_deaths, 10)
```

```{r}
tail(all_deaths, 1)
```

```{r}
glimpse(all_deaths)
```
```{r}
summary(all_deaths)
```

```{r}
str(all_deaths)
```

#Opens in separate window like Excel
```{r}
view(all_deaths)
```

```{r}
test <- all_deaths
```

```{r}
select(all_deaths, name, allegiances, gender, nobility, year_of_death)
```

```{r}
deaths_no_name <- select(all_deaths, -name)
```

```{r}
deaths_no_name
```

Select all columns which relate to a book
Select the columns which are related to some kind of death

```{r}
select(all_deaths, contains("book"))
```
```{r}
select(all_deaths, contains("death"))
```

```{r}
select(all_deaths, name:book_of_death)
```

```{r}
all_deaths
```

== tests if two things are equal.
!= tests if two things are not equal.
< tests if the left is less than the right.
> tests if the left is greater than the right.
<= tests if the left is less than or equal to the right, >= tests for greater than or equal to.
%in% tests if the left is contained in the right.

```{r}
filter(all_deaths, allegiances != "Lannister")
```

```{r}
filter(all_deaths, allegiances %in% c("Lannister", "House Lannister"))
```

```{r}
filter(all_deaths, year_of_death >= 299)
```

```{r}
filter(all_deaths, allegiances == "None")
```
& and
| or
```{r}
filter(all_deaths, allegiances == "Lannister" & nobility == 1)
```

```{r}
filter(all_deaths, allegiances == "Lannister" | nobility ==1)
```

```{r}
filter(all_deaths, 
       book_intro_chapter >= 5 & book_intro_chapter <= 10)
```

```{r}
filter(all_deaths, between(book_intro_chapter, 5, 10))
```

```{r}
filter(all_deaths, str_detect(allegiances, "Lannister"))
```

Find where the year_of_death is less than or equal to 299
Find the females(gender is 0) who are not Lannisters
Find just the data for the characters "Jon Snow", "Daenerys Targaryen" and "Samwell Tarley"

```{r}
filter(all_deaths, year_of_death <= 299)
```

```{r}
filter(all_deaths, gender == 0 & allegiances != "Lannister")
```

```{r}
filter(all_deaths, 
       name == "Jon Snow" | 
         name == "Daenerys Targaryen" | 
         name == "Samwell Tarly")
```

```{r}
filter(all_deaths, name %in% c("Jon Snow", "Daenerys Targaryen", "Samwell Tarly"))
```

```{r}
arrange(all_deaths, book_intro_chapter)
```

```{r}
arrange(all_deaths, desc(book_intro_chapter))
```

```{r}
arrange(all_deaths, book_of_death, death_chapter)
```

Arrange all deaths by allegiances. 
What happens when you arrange by a character column?
It puts them in alphabetical order
Arrange all deaths by allegiances and book intro chapter
Arrange all deaths by descending year of death

```{r}
arrange(all_deaths, allegiances)
```

```{r}
arrange(all_deaths, allegiances, book_intro_chapter)
```

```{r}
arrange(all_deaths, desc(year_of_death))
```

```{r}
mutate(all_deaths, years_survived = year_of_death - 298, 
       .after = year_of_death)
```

```{r}
mutate(all_deaths, 
       test = "test")
```

```{r}
mutate(all_deaths, 
       year_of_death = is.na(year_of_death))
```

```{r}
mutate(all_deaths, 
       nobility = as.logical(nobility))
```

```{r}
mutate(all_deaths, 
       avg_death = mean(year_of_death, na.rm = TRUE))
```

```{r}
summarise(all_deaths, 
          avg_death = mean(year_of_death, na.rm = TRUE), 
          sum_males = sum(gender))
```

```{r}
deaths_grouped <- group_by(all_deaths, allegiances)
```

```{r}
summarise(deaths_grouped,
          character_count = n())
```

```{r}
deaths_grouped <- group_by(all_deaths, nobility, gender)
```

```{r}
summarise(deaths_grouped,
          char_count = n())
```

```{r}
have_died <- filter(all_deaths, !is.na(book_of_death)) 

died_grouped <- group_by(have_died, allegiances)
died_counts <- summarise(died_grouped, count = n())

arrange(died_counts, desc(count))
```

```{r}
all_deaths %>% 
  filter(!is.na(book_of_death)) %>% 
  group_by(allegiances) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```rt

