---
title: "R Notebook"
output: html_notebook
---

# Connecting R to a Database
- Get data from a database and bring into R for further analysis
- Connect R to postgres database (omni pool)

## Don't put login credentials on github (or any public place)

- be careful with how we work with our login details

-- host name
-- username
-- database
-- port
-- password

Strategies for protecting login details
- gitignore
-store outside of a git repo

Load in our secret login credentials into R

```{r}
source("secrets.R")
```



Connecting to a database

```{r}
library(RPostgres)

db_connection <- dbConnect(
  drv = Postgres(),
  host = om_host,
  port = om_port,
  dbname = om_atabase,
  user = om_username,
  password = om_password,
  bigint = "numeric"
)

# not that we've created our connection object, we don't need our credentials anymore - we can remove these from our environment

rm(om_username,
   om_password)

```


## Testing our connection


```{r}
dbListTables (conn = db_connection)
```


## Running queries

```{r}
employees_table <- dbGetQuery(conn = db_connection, statement = "SELECT * FROM employees;")
```

```{r}
library (tidyverse)
```

```{r}
employees_table %>% 
  filter(department == "Legal")
```


## Using SQL code chunks in R Notebooks


#To put SQL code into R, you need to specify the language in the code chunk
and then give it a connection
```{sql, connection = db_connection}
SELECT *
FROM employees
LIMIT 10
```


Running a query and capturing the output
creates/updates an object. In this case "out"


```{sql, connection = db_connection, output.var = "out"}
SELECT *
FROM employees
LIMIT 10
```

Then we can call "out" in R
```{r}
out %>% 
  select(first_name, last_name, department)
```

## When finished querying the database, you should disconnect 

```{r}

```

