---
title: "R Notebook"
output: html_notebook
---

## Learning objectives

- Know how to create a basic Shiny app
-Understand how to use inputs and outputs to create interactively

## Data

```{r}
library(tidyverse)
```


```{r}
olympic_medals <- read_csv("data/olympics_overall_medals.csv")
```

Jump to - Ctrl + P
Ctrl + i - reindents when you've messed up the indentation
Ctrl + a - highlights everything

## Widgets

Needs: ID, Label, choices

```{r}
widgetInput(inputID = "id",
            label = "widget label",
            choices = c("choice A", "choice B", "choice C"))
```

render function --> output function


## Plot

```{r}
olympic_overall_medals %>%
  filter(team == "Great Britain",
         season == "Summer") %>%
ggplot(aes(x = medal, y = count, fill = medal)) +
  geom_col()
```

## Learning Objectives

- Know how to include HTML in a Shiny app
- Know how to include CSS in a Shiny app
- Be able to use themes and UI Layouts in Shiny

Using HTML in Shiny

tags$
HTML()

## Theme-ing packages

"{bslib}" - Bootswatch library

Each row has a width of 12, so the max value a column within that row can take is 12. If you want 4 columns of equal width, for example, you woudl make each 3

There are no spaces between columns in a row, but you can offset them