---
title: "R Notebook"
output: html_notebook
---
```{r}

```

##Session 1 -- Using dates and times

## Learning objectives

- Be able to manipulate dates and times in R
- Understand about time zones

iso 8601 standard - international standard for dealing with dates and times


Task - 5 mins - Take a few min and read about ISO 8601 standard, as we will be working with it a lot today. Answer the following questions:

- What order does ISO standard follow
YYYY-MM-DD
- What is the minimum number of digits you can have in a year? Why?
4 digits - to avoid the year 2000 problem (potential computer errors relating to the formatting and storage of calendar data that might come from only storing the final two digits)
-What clock system do the times use?
24 hour clock
- How many time formats does ISO standard have
4?
- What timezone are time and dates automatically represented in?
UTC - Universal coordinated time
- How are combined date-times represented?
You use the letter T as a delimiter, e.g. 2007-04-05T14:30

YYY = four digit year
MM = two digit month
DD = Two digit day of the month
hh - two digits of hour in 24 hour time
mm = two digits of minute
ss = two digits of seconds
s - one or more digits 
TZD

```{r}
as.Date("2022-08-21")
```

Dates are represented as integer numbers, so if you unclass them, it will tell you the number of days

```{r}
date_time1 <- as.POSIXct("2020-01-01 00:00:00")
date_time1
```


```{r}
date_time2 <- as.POSIXlt("2020-01-01 00:00:00")
date_time2
unclass(date_time2)

date_time2$mday
```

### Packages

```{r}
library(tidyverse)
library(lubridate)
```
Date conversion functions like ymd()

Date-time converting functions
ymd_hms()

Extracting components
year()
wday()
hour()
etc.

timespans
as.period()






```{r}
netflix <- read_csv("data/netflix.csv")
```

Date: "<date>"
Time: "<time>"


```{r}
"2019-01-19"

ymd("2019-01-19")

mdy("January 9th, 2019")

dmy("9-Jan-2019")

ymd_hms("2019-01-19 10:08:26")
ymd_hm("2019-01-19 10:08")
ymd_h("2019-01-19 10")
```
Change column date added from character to a date using either dmy mdy or ymd

```{r}
netflix <- netflix %>% 
  mutate(date_added = mdy(date_added))
```

## Extracting individual components

```{r}
month(netflix$date_added[1:5], label = TRUE, abbr = FALSE)

netflix <- netflix %>% 
  mutate(day = day(date_added),
         month = month(date_added, label = TRUE, abbr = FALSE),
         year = year(date_added)
         )
```

Add a column weekday which is the day of the week thhe title was added

```{r}
netflix <- netflix %>% 
  mutate(weekday = wday(date_added, label = TRUE))
```
 
 ## Calculating time spans
 
 Three types of time span
 
 Difftime: a datetime object that represents days 
 Period: Represent a timespan like week/month/day - a discreet period
 Duration: represent an exact number of seconds
 
 today() <- gives today's date
 now() <- gives today's date and the time inc seconds
 
```{r}
netflix<- netflix %>% 
  mutate(time_difftime = today() - date_added)
```
 
```{r}
netflix <- netflix %>% 
  mutate(time_period = as.period(today() - date_added),
         time_years = as.numeric(time_period, "years"),
         time_hours = as.numeric(time_period, "hours"),
         time_minutes = as.numeric(time_period, "minutes"),
         time_seconds = as.numeric(time_period, "seconds"))
netflix
```
 
### Subtracting periods from dates

For example, say we found out that Netflix promotes releases on their web app for 2 weeks before the show is released and then takes them down after 4 months

```{r}
netflix <- netflix %>% 
  mutate(promo_date = date_added - weeks(2),
         removal_date = date_added +months(4))

netflix %>% 
  select(date_added, promo_date, removal_date)
```

```{r}
netflix %>% 
  mutate(time_period = as.period(today() - date_added),
         weeks = as.numeric(time_period, "weeks"))
```

Duration: represent an exact number of seconds

```{r}
netflix <- netflix %>% 
  mutate(time_duration = as.duration(today() - date_added))

netflix %>% 
  select(starts_with("time"))
```

dseconds()
dminutes()
dhours()
ddays()
dweeks()
dmonths()
dyears()

What date is a year after the 10th of January 2020?

Is it:

```{r}
ymd(20200110) + years(1)
ymd(20200110) + dyears(1) #adds on 365 days exactly, so one day less b/c of leap year in 2020
```

## Understanding timezones

```{r}
OlsonNames()

ymd_hms("2023-09-13 10:22:01", tz = "Atlantic/Faeroe")


netflix %>% 
  mutate(date_tz_1 = force_tz(date_added, tzone = "Asia/Singapore"),
         date_tz_2 = with_tz(date_added, tzone = "Asia/Singapore")) %>% 
  select(date_added, date_tz_1, date_tz_2)
```

What is the difference between the two functions?

force_tz just adds a time zone, it doesn't change the date or time.
with_tz figures out what the date and time would be if you moved from the current time zone to the one asked for


## Session 3 -- Introduction to time series

### Learning objectives
-Know what time series data looks like
-Understand some of the key components of time series data
-See some examples of the application of time series data
-Understand why familiarity with time series data is important

```{r}
library(tsibble)
library(tsibbledata)
```

"as_tsibble()"

## "tsibble"

__index__: time component
__Key__  : identifies the underlying structure of the data
         : in conjuction with the __index__ uniquely identities each individual row
         

```{r}
vic_elec
```

```{r}
index(vic_elec)
index_var(vic_elec)

key(vic_elec)
key_vars(vic_elec)

interval(vic_elec)
```

```{r}
vic_elec %>% 
  select(Time, Demand, Temperature)
```

```{r}
vic_elec %>% 
  filter(!Holiday)

vic_elec %>% 
  filter(Time < "2012-01-01 01:00:00")

vic_elec %>% 
  filter(Time < as.POSIXct("2012-01-01 01:00:00", tz = "Australia/Melbourne"))

vic_elec %>% 
  filter(Time < as.Date("2012-01-02"))

vic_elec %>% 
  filter_index("2012-01")
```




```{r}
vic_elec <- vic_elec %>% 
  mutate(year = lubridate::year(Date))
```

```{r}
vic_elec %>% 
  select(year, Temperature) %>% 
  group_by(year) %>% 
  summarise(mean_temp = mean(Temperature))

vic_elec %>% 
  select(year, Temperature) %>% 
  index_by(year) %>% 
  summarise(mean_temp = mean(Temperature))

vic_elec %>% 
  mutate(date = as_date(Time), .after = Time)


vic_elec %>% 
  index_by(as_date(Time)) %>% 
  summarise(mean_temp = mean(Temperature))
```

## "tsibble" and "ggplot2"

```{r}
vic_elec %>% 
  filter(!Holiday) %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Temperature))
```
Useful for extractign different components of the time series data
```{r}
floor_date()
ceiling_date()

yearweek()
yearmonth()
yearquarter()
```

```{r}
elec_date <- vic_elec %>% 
  index_by(date = as_date(Time)) %>% 
  summarise(temp_mean = mean(Temperature, na.rm = TRUE))

elec_date %>% 
  ggplot(aes(date, temp_mean)) +
  geom_line()

elec_month <- vic_elec %>% 
  index_by(month = month(Time, label = TRUE)) %>% 
  summarise(temp_mean = mean(Temperature, na.rm = TRUE))

elec_month %>% 
  ggplot(aes(month, temp_mean)) +
  geom_point() +
  geom_line(group = 1)
```

```{r}
elec_year <- vic_elec %>%
  index_by(year = year(Time)) %>%
  summarise(temp_mean = mean(Temperature, na.rm = TRUE))

elec_year

elec_year %>%
  ggplot(aes(x = year, y = temp_mean)) +
  geom_col(fill = "steelblue", alpha = 0.7) + 
  ylab("Mean Temperature")+ 
  xlab("year")
```


## Rolling windows


```{r}
library(slider)
library(feasts)
```

```{r}
elec_rolling <- vic_elec %>% 
  mutate(
    temp_moving_avg = slide_dbl(
      .x = Temperature,
      .f = ~ mean(., na.rm = TRUE),
      .before = 1000,
      .after = 1000,
      .complete = TRUE
    ),
    temp_mvn_avg_1c = slide_dbl(
      .x = Temperature,
      .f = ~ mean(., na.rm = TRUE),
      .before = 100,
      .after = 100
    ),
    temp_mvn_avg_comp = slide_dbl(
       .x = Temperature,
      .f = ~ mean(., na.rm = TRUE),
      .before = 2000,
      .after = 2000,
      .complete = FALSE
    )
  )

ggplot(elec_rolling) +
  geom_line(aes(Date, Temperature), colour = "grey") +
geom_line(aes(Date, temp_moving_avg), colour = "red") +
geom_line(aes(Date, temp_mvn_avg_1c), colour = "blue") +
geom_line(aes(Date, temp_mvn_avg_comp), colour = "cyan")
```

```{r}
ggplot(elec_rolling) +
  geom_line(aes(x = Date,
                y = Temperature), colour = "grey") +
  geom_line(aes(x = Date,
                y = temp_moving_avg), colour = "red")


```


## "feasts"


```{r}
tourism

index_var(tourism)
key_vars(tourism)


tourism %>% 
  distinct(Purpose)

tourism %>% 
  distinct(State)

tourism %>% 
  distinct(Region)

tourism %>% 
  distinct(State, Region, Purpose)
```

```{r}
holidays <- tourism %>% 
  filter(Purpose %in% c("Holiday", "Business")) %>% 
  group_by(State) %>% 
  summarise(Trips = sum(Trips))

holidays

holidays %>% 
autoplot(Trips) +
  xlab("Year")
```

```{r}
holidays %>% 
  filter(State %in% c("Queensland", "New South Wales", "Victoria")) %>% 
  gg_season(Trips)
```

```{r}
holidays %>% 
  filter(State %in% c("Queensland", "New South Wales", "Victoria")) %>% 
  gg_subseries(Trips)
```

Do the hard work for your audience
And make it nice for the audience