---
title: "R Notebook"
output: html_notebook
---

# Today's agenda

More of a "programmy" day
Can sometimes feel divorced from analysis techniques as we've learned so far

1. Data structures
2. Control flow
3. Functions

# Session 1 - Matrices and Lists

## Learning objectives

-Know what lists and matrices are
  -Matrix is a 2D structure (like a dataframe)
  -List is a non-atomic data structure - can contain several values and data types
-Know how to create lists and matrices
  -list()
  -matrix()
-Be able to do some basic manipulations on lists and matrices
  -"[[]]"
  -"[]"
  -for lists: "$", remember that a data frame is a list, so can also use "$" in data frames
  -for matrices - index by _row_, then _column_

# Lists

```{r}
x <- list(1, "a", TRUE, c(1, 2, 3))

v <- 1:5

c(1, "a", TRUE, c(1, 2, 3))

as.list(v)
```

## Subset

```{r}
#Need double square brackets when trying to return a vector from a list - if you use single square brakets, it will return a list
x[[2]]

x[2]
```

##Named list

```{r}
named_list <- list(a = "I love R", b = 1:10)
named_list$a
```

-Data frames are lists.
-Some tidyverse functions work the same as base R functions - e.g. pull() function works the same as double square brackets [[]]
-The best argument for tidyverse is that it shows a clear progression - it is clear what is happening

```{r}
student_info <- list(
                Agnes = list(height = 1.8, shoe_size = 8L),
                Bobby = list(height = 2.1, gender = "Male"),
                Charlie = list(gender = "Female", likes_beets = TRUE)
)

str(student_info)
n <- "Agnes"
student_info$Agnes$height #Use this if you are sure which column you want - and it won't CHANGE
student_info[["Agnes"]][["height"]]#Use this if you want to access programmatically, also because its safer
```

## Task
Make a named list that contains a numeric vector and a character vector.
    Access the second element of your list using double square brackets [[]]. What is it’s class?
    Access the second element of your list using single square brackets []. What is it’s class?
    Access the second element of your list by name. What is it’s class?
    
```{r}
task_list<- list(a = "Hello all", b = 1:5)

task_list[[2]] #class integer

task_list[2] #class list

task_list$b #class integer
```

c() - combines elements together into a "flat" structure e.g. a vector
list() - combines elements together 

# Matrices

Data frames are heterogeneous data structures - multiple data classes are allowed
Matrices are homogenous data structures - only one data class is allowed.

## Creating matrices
```{r}
m <- matrix (1:12, ncol = 3, byrow = FALSE) # R uses column order (e.g. fills down columns first) unless you use the byrow argument

class(m)
typeof(m)
```

## Indexing a matrix

m[row number, column number]

```{r}

m[2:3, ]
m
m[2:3, 2]
m[c(1, 3), c(1, 3)]
m[c(1, 3), -2]
```
Matrices work in a vectorized way
```{r}
v * 2

m - 2
```

```{r}
rowSums(m)
colSums(m)

rowMeans(m)
colMeans(m)
```

## Task

Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 - 3

```{r}
task_matrix <- matrix(1:3, nrow = 3, ncol = 4)
```



Find the element in the 2nd row, 3rd column
Check that all the columns are identical by checking that their sums are the same
```{r}
#Select the element in the 2nd row, 3rd column
task_matrix[2, 3]
#Take sum of each column
colSums(task_matrix)
```

```{r}
rep(1:3, each = 3)
colnames(m) <- paste0("col_", 1:3)
rownames(m) <- paste0("row_", 1:4) #paste0 function doesn't have a default seperator, paste has one space as a default seperator.
```

#Session 2 -- Control flow

##Learning objectives
-Know the structure of all the conditional statements
-Be able to nest conditional statements
-Know how to use "for" loops, but understand when they should and should not be used

#"if" statements

if(this thing is true){# Then
Do this thing
}

if(weather == "Sunny"){
hang_up_washing()
}


```{r}
x <- 10
y <- 15

if(x < y){#then
  print("yes")}

```
If is not a vectorized element, so it doesn't recycle over an entire vector. For example, the below will return an error unless the logical operator is changed to %in%
```{r}
x <- 10
y <- 5:20

if(x < y){#then
  print("yes")}

```

## Task

```{r}

x <- 21
  
if(x %% 2 == 0){
  print("Yes that divides by 2")
}
```

#If else statements

if(this thing is true){# if the condition is TRUE, Then
Do this thing
}else{ #otherwise,
Do this thing instead
}

if(weather == "Rainy"){
wear_wellies()
} else {
wear_trainers()
}

```{r}

x <- 21
y <- 10:50
  
if(x %% 2 == 0){
  print("Yes that divides by 2")
} else {
  print("No")
}

if(x %in% y){
  print("Yes")
} else {
  print("No")
}
```

## Task - 1 - Edit your code from the first exercise so that it prints 'Yes that divides' when x is divisible by 2, and prints "No that does not divide"

```{r}
x <- 21
y <- 10:50
  
if(x %% 2 == 0){
  print("Yes that divides by 2")
} else {
  print("No that does not divide")
}
```
## Task 2 - 
Here's an example of a nested if-else statement.
Predict what it's going to print out for the following conditions.
-When x is 5 and y is 10
-When x is 6 and y if 12
-When x is 6 and y is 13

```{r}
x <- 6
y <- 13

if (x %in% c(3, 4, 5)){
  
  print("x is in our list")
  
  if (x == 5){
    
    print("x is exactly 5")
    
  }
  
} else {
  
  print("Start of else statement")
  
  if (y %% x == 0){
    
    print("x divides y")
  
  } else{
    
    print("x does not divide Y")
  }
  
  print ("End of else statement")
  
}
```
# Loops

- "for" loops 
  - used when we know how many iterations we need
  - ends after desired number of loops
  
- "while" loops 
  - used when we don't know how many iterations we need
  - will end when a condition is met

Bear in mind:
Loops are dead useful,
BUT we usually can avoid them in R

## for() loops

For loops might be slow, but probably because a failure to vectorize

for (element in vector){
print(element * 2)
}


```{r}
v <- 10:20

v * 2
```
```{r}
for (item in v) {
 print(item * 2) 
}
```

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (day in weekdays){
  #phrase <- paste("The day is", day)
  #print(phrase)
  print(paste("The day is", day))#short version, use long version for more complicated
}
```

## while() loops

Careful - these can sometimes get stuck in an infinite loop

```{r}
current_value <- 1 #initialise your variable first
stop_value <- 10

while(current_value < stop_value){
  current_value <- current_value + 1
  print(current_value)
}
```

## Alternative ways to index for a loop

```{r}
weekdays
len_weekdays <- length(weekdays)

for(i in 1:len_weekdays){
  phrase <- paste("The day is", weekdays[i])
  print(phrase)
}
```

## Replacing a vcetor with a loop __TERRIBLE PRACTICE__

```{r}
weekdays
len_weekdays <- length(weekdays)
for(i in 1:len_weekdays){
  plural <- paste0(weekdays[i], "s") #creating a new value from the old based on the index i
  weekdays[i] <- plural #modifying each vector element based on the index i
}

```
The above is a very complicated way of doing this:

```{r}
paste0(weekdays, "s")
```

## Nested loops

```{r}
m <- matrix(1:12, 4)
nrows <- nrow(m)
ncols <- ncol(m)

for(row_i in 1:nrows){
  print(m[row_i, ])
}

for(row_i in 1:nrows){
  for(col_i in 1:ncols){
    print(m[row_i, col_i])
  }
}
```

```{r}
product_matrix <- matrix(NA, nrow = 5, ncol = 5)
nrows <- dim(product_matrix)[1]
ncols <- dim(product_matrix)[2]

for(i in 1: nrows(product_matrix)){
  for (j in 1:ncols(product_matrix))
  #
  product_matrix[i, j] <- i * j
}
```

```{r}
example_matrix = matrix(1:25, nrow = 5, ncol = 5)
example_matrix 

for (row_index in 1:5){
  for (col_index in 1:5){
    if (row_index == col_index){
      print(example_matrix[row_index, col_index])
    }
   }
}
```

```{r}
paste0("Today is ", weekdays)
```

# Functions
## Learning objectives

-Be able to write your own functions
-Know best practice for writing R functions

R is mostly a functional programming language
__OOP__ object oriented programming language

A function is an isolated block of code that accepts an input (or inputs) and returns an output.

Think of it as being like a machine on a production line: 
unprocessed items go into the machine, processed items come out

'nchar' 
'nchar(x, type = "chars", allowNA = FALSE, keepNA = NA)
-Arguments need to be in the order they are listed unless you're going to name them explicitly

parameter <---> argument (parameter and argument are interchangable)
```{r}
nchar("hello", allowNA = TRUE)
```

'paste' (..., sep = " ", collapse = NULL, recycle0 = FALSE)
Any parameters after dots must be named
```{r}
paste("a", "b", "c", "d", sep = "--")
```

## Writing our own functions

```{r}
function_name <- function(par_1, par_2 = NULL, par_3 = "default value")){
  #stuff to to do
  
  #the last thing that's done will be the return value
}
```

```{r}
multiply_by_5 <- function(x) {
  x * 5
}
multiply_by_5(2)
```

```{r}
multiply_by <- function(x, y) {
  z <- x * y
  z
}

multiply_by(2, 10)
```

```{r}
multiply_by <- function(x, y = 3) {
  z <- x * y
  z
}
```

```{r}
make_price <- function(x, currency = "£"){
  paste0(currency, x)
}

make_price(3)
```



## Scope

What can a function "see"?

```{r}
n <- 8
multiply_by_n <- function(x) {
  x * n
}
```

Global scope - variables specified in the R Global environment
Local scope - local to the function

Function can look into the global environment
Don't use variables inside the function that aren't specified in the function b/c it will look outside the local scope if it can't find and use variables from the global scope (some of which you might not be aware of.)

```{r}
multiply_by_n <- function(x, n){
  ans <- x * n
  if(ans %% 2 == 1){
    z <- paste("x is odd")
  }else{
    z <- paste("x is even")
  }
  z
}
multiply_by_n(3, 3)
```

## Best practice for writing functions

__good__

Name with verbs if you can

"add_3()"
"format_number()"
"read_data()"

__bad__

"adder()"
"number_formatter()"
"data_reader()"

-Braces {} should start on a line and then end on their own line - for example:
```{r}
n <- 8
multiply_by_n <- function(x) {
  x * n
}
```

__DRY__ principle
Don't Repeat Yourself

## Debugging

debugonce()