---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```
# Lab
## Q1: Are the Thought Police ever mentioned?



```{r}
#Detect thought police, ignore case
whole_text %>% 
  str_detect("(?i)Thought Police")
```
## Q2: Count how often big brother is mentioned

```{r}
#Counts the number of instances of big brother, ignore case
whole_text %>% 
  str_count("(?i)Big Brother")
```

## Q3: The chapter is written from the point of view of the protagonist, Winston. Does he ever mention the other main character (Julia) by name?
No

```{r}
#Detect Julia, ignore case
whole_text %>% 
  str_detect("(?i)Julia")
```

## Q4: Winston writes the date in a specific format: month nth, year eg. August 5th, 2019. Write a regular expression to match dates in that format and find the date written in his diary.

```{r}
whole_text %>% 
  str_extract_all("[A-Z][a-z]+[:space:][1-9]{1,2}[a-z]{2}[:punct:][:space:][1-9]{4}")
```
##Q5: 

```{r}
whole_text %>% 
  str_replace_all("Winston | Winston Smith", "REDACTED")
```

Q6

```{r}
whole_text %>% 
  str_extract_all("(?i)[:space:]gin[[:space:][:punct:]]")
```
```{r}
whole_text <- whole_text %>% 
  str_replace_all(c(" GIN." = " VODKA.",
                    " gin." = " vodka.",
                    " gin " = " vodka "))
```

