---
title: "R Notebook"
output: html_notebook
---

# Libraries
```{r}
library(tidyverse)
```


# Lesson 1 - Strings

```{r}
single <- 'this is a single string'
double <- "this is a double string"
```

```{r}
single
double
```


```{r}
employees <- tibble(
  first_name = c("Sally", "steven", "Mary", "John"),
  last_name = c("Smith", "hasTings", "Milne", "Davidson"),
  postcode = c("EH10 6GH", "EH3 7HL", "G1 7FR", NA), 
  location = c(" ", "Edinburgh", "", "Stirling")
)

employees
```

```{r}
employees %>% 
  group_by(location) %>% 
  summarise(n())
```

```{r}
mary_location <- employees %>% 
  filter(first_name == "Mary") %>% 
  select(location) %>% 
  pull()

sally_location <- employees %>% 
  filter(first_name == "Sally") %>% 
  select(location) %>% 
  pull()
```

```{r}
mary_location == sally_location
```
```{r}
employees %>% 
  summarise(nchar(location))
```

```{r}
str_c("This", "is", "some", "text", sep = " ")
```

```{r}
employees %>% 
  mutate(full_name = str_c(first_name, last_name, sep = " "),
         .after = last_name)
```

```{r}
employees %>% 
  mutate(full_name = paste(first_name, postcode, sep = " "),
         .after = last_name)
```

The paste function in base R will change NAs into a string when joining while the str_c function won't join an NA to a string.

```{r}
pangram <- "The quick brown fox jumps over the lazy dog"
```

```{r}
pangram %>% 
  str_to_upper()
```

```{r}
pangram %>% 
  str_to_lower()
```

```{r}
pangram %>% 
  str_to_title()
```

```{r}
pangram %>% 
  str_to_sentence()
```

```{r}
employees %>% 
  select(first_name, last_name) %>% 
  mutate(first_name_fix = str_to_title(first_name),
         last_name_fix = str_to_title(last_name))
```

```{r}
unbroken_string <- "We're going to break up this string."
```

```{r}
unbroken_string %>% 
  str_split(" ") %>% 
  flatten_chr()
```

```{r}
employees %>% 
  mutate(new_postcode = str_split(postcode, " ")) %>% 
  #takes strings out of list and gives each one their own row.
  unnest(new_postcode)
```


```{r}
employees %>% 
  separate_rows(postcode, sep = " ")
```

```{r}
employees %>% 
  separate(postcode, c("postcode_outer", "postcode_inner"), sep = " ")
```

## Task - 10 mins
Concatenate the strings "Manipulating" "strings" "is" "fun!"
Capitalise the result and save it in a new variable
Split the formatted string into individual characters

```{r}
loud_string <- str_c("Manipulating", "strings", "is", "fun!", sep = " ") %>% 
  str_to_upper()

loud_string %>% 
  str_split("") %>% 
  flatten_chr()
```

## More strings

```{r}
full_string <- "Let's grab a substring from this!"
```

```{r}
full_string %>% 
  str_sub(start = 6)
```
```{r}
full_string %>% 
  str_sub(-5, -1)
```

```{r}
employees %>% 
  mutate(postcode_inner = str_sub(postcode, -3),
         postcode_outer = str_sub(postcode, 1, -4))
```

```{r}
string <- "We're going to replace some words."
```

```{r}
str_sub(string, 1, 22) <- "I've changed"
```

```{r}
number <- 123456.789
```

```{r}
number %>% 
  format(nsmall = 1, decimal.mark = ",")
```

```{r}
number %>% 
  format(big.mark = ",", big.interval = 3) %>% 
```
When commas are introduced into a string that is a number, it can't be re-rendered as numeric using as.numeric()

## Task - 10 mins
```{r}
spotify_songs <- read_csv("spotify_sample.csv")
```


-Upcase the text in the playlist_name column. Overwrite the existing column to store the result.

```{r}
spotify_songs <- spotify_songs %>% 
  mutate(playlist_name = str_to_upper(playlist_name))
```

-Lowcase the text in the track_artist column. Store the result in a new column called track_artist_lowcase.

```{r}
spotify_songs <- spotify_songs %>% 
  mutate(track_artist_lowcase = str_to_lower(track_artist), 
         .after = track_artist)
spotify_songs
```


Take the first 5 characters of the playlist_id column, and save it in a new column called playlist_start.

```{r}
spotify_songs %>% 
  mutate(playlist_start = str_sub(playlist_id, start = 1, end = 5))
```


Take the first 4 characters of the track_album_release_date column and save it in a new column called year.

```{r}
spotify_songs %>% 
  mutate(year = str_sub(track_album_release_date, 1, 4))
```


What else can you do to this dataset to practice? Can you work with the numbers? This is open ended - have fun!

#Lesson 2 - Regular expressions

Learning outcomes:
What is a regular expressions
Use them to match patterns in text
Create regular expressions

```{r}
single_string <- "string a"
strings <- c("string a", "string b", "string c", "string d", "striiiiiiing e")
```

```{r}
pattern <- "a"
```

```{r}
single_string %>% 
  str_detect(pattern, negate = TRUE)
```
```{r}
strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "str"

strings %>% 
  str_detect(pattern)
```

```{r}
#square brackets cause R to search for a OR b OR c - so will return TRUE if any of these are present in the string
pattern <- "[a-z]"

strings %>% 
  str_detect(pattern)
```

Regular expressions are case sensitive by nature, you need to add (?i) to make it ignore the case

```{r}
#adding (?i) causes R to ignore the case
pattern <- "(?i)[A-Z]"

strings %>% 
  str_detect(pattern)
```

```{r}
#Using the curly brackets specifies the number of the chosen character you're looking for
pattern <- "i{3}"

strings %>% 
  str_detect(pattern)
```

```{r}
#Adding the comma after the number in the curly bracket allows you to look for n or more (3 or more)
pattern <- "i{3,}"
#Find between 3 and 6 i's
pattern <- "i{3,6}"

strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "str|gr"

strings %>% 
  str_detect(pattern)
```
## Task - 15 min

```{r}
codeclan_strings <- c(
  "We're having great fun learning about data analysis at CodeClan!",
"CodeClan is at 37 Castle Terrace in Edinburgh and ClockWise in Glasgow",
"Our cohort code starts with DE or DR",
"You get a great view of the castle or the national conservatoire from just outside CodeClan's door")
```

```{r}
codeclan_strings %>% 
  str_detect("CodeClan")
```

```{r}
codeclan_strings %>% 
  str_detect("[:upper:]{2}")

codeclan_strings %>% 
  str_detect("[A-Z][A-Z]")
```

```{r}
pattern <- "(?i)Castle"

codeclan_strings %>% 
  str_detect(pattern, negate = TRUE)
```


## regex continued

```{r}
strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)
```


## Useful operators in regex
? - zero or one {0,1}
* - zero or more {0,}
+ - one or more {1,}

```{r}
string <- "@emailprovider.com"
pattern <- "[a-z]+@emailprovider.com"

string %>% 
  str_detect(pattern)
```


```{r}
pattern <- "[a-z0-9[:punct:]]+@[a-z[:punct:]]+"

strings %>% 
  str_extract_all(pattern)
```

When you want to detect ".", you need to add two backslashes to escape their special meaning "\\."

```{r}
pattern <- "[a-z0-9[:punct:]]+@[a-z[:punct:]]+"

strings %>% 
  str_replace_all(pattern, "REDACTED")
```

## Task - 15 min
- Make a vector containing the strings:
"Colin's mobile number definitely isn't 07777777777."
"A landline might be 01010 123456."
"Mhairi, Del, Aileen and Steph helped write these notes"
"I'm called your_name_here and I live in your_town_here."
Use regular expressions to:
-Extract all the names
-Extract the landline number
-Insert your name and where you live into the fourth string

```{r}
contact_string <- c(
  "Colin's mobile number definitely isn't 07777777777.",
  "A landline might be 01010 123456.",
  "Mhairi, Del, Aileen and Steph helped write these notes",
  "I'm called your_name_here and I live in your_town_here."
  )
```

```{r}
#Extract all the names
contact_string %>% 
str_extract_all("[A-Z][a-z]+")
```

```{r}
#Extract the landline number
contact_string %>% 
str_extract_all("[:digit:]+[:space:][:digit:]+")
```

```{r}
#insert your name and where you live
pattern1 <- "[a-z_]+name+[a-z_]+"
pattern2 <- "[a-z_]+town+[a-z_]+"

contact_string %>% 
  str_replace_all(c(pattern1, pattern2), c("Jamie", "Glasgow"))
```


```{r}
test_strings <- c(
  "abc", 
  "Abc",
  "abC",
  "s"
)
```

```{r}
pattern <- "^[a-z]+[a-zA-z]*"
pattern2 <- "[a-zA-z]*[a-z]$"
pattern3 <- "^[a-z]+$"

test_strings %>% 
  str_extract_all(pattern3)
```

```{r}
student <- c("Abigail", "Bjorn", "Caitlin")
age <- c(25, 56, 40)
emails <- c("abigal12@company.co.uk", "bsmith@gmail.co.uk", "caitlin_1980")

student_info <- tibble(student, age, emails)

student_info 
```


```{r}
student_info %>% 
  mutate(email_detect = str_detect(emails, "[a-z0-9]+@"))
```

```{r}
student_info %>% 
  mutate(emails = str_extract(emails, "@[a-z.]+"))
```

