---
title: "R Notebook"
output: html_notebook
---
# Libraries

```{r}
library(tidyverse)
```


## Session 1 -- Core definitions

### Learning objectives recap

- Recognise what is meant by the terms probability and statistics
- Gain an understanding of core probability concepts
- Gain an awareness of the mathematical properties of a probability


```{r}
source("prob.R")
```

```{r}
s_two_dice <- rolldie(2)
```

```{r}
s_two_dice
s_two_dice %>% 
  filter(X1 + X2 == 7) 
```

```{r}
s_two_dice <- s_two_dice %>% 
  mutate(total = X1 + X2)
```

```{r}
s_sum_equals_seven <- s_two_dice %>% 
  filter(total == 7)

s_sum_gt_10 <- s_two_dice %>% 
  filter(total > 10)
```
###Task
```{r}
first_die_even <- s_two_dice %>% 
  filter(X1 %% 2 == 0)

first_even_second_odd <- s_two_dice %>% 
  filter(X1 %% 2 == 0,
         X2 %% 2 == TRUE)
```

## Session 2 -- Counting rules

### Learning objectives

- Understand why _counting rules_ are important in probability theory
- Know the counting rules for the following cases:
    -single step experiments
    -multiple step experiments
    -permutations
    -combinations
    
### Single step experiments

example: throwing two dice in one go

$$
prob(event) = \frac{num(event\ outcomes)}{num(all\ outcomes)}
$$



```{r}
s <- rolldie(2)

total_seven <- s %>% 
  filter(X1 + X2 == 7)
```

```{r}
nrow(total_seven) / nrow(s)
```

### Multiple step experiments

Step 1: rolling a dice - 6 possible outcomes
Step 2: tossing a coin - 2 possible outcomes

6 * 2

What is the probability of 

P(H) = 1/2
P(6) = 1/6

P(H) & P(6) = 1/2 * 1/6 = 1/12


```{r}
s_die <- rolldie(1)
s_coin <- tosscoin(1)
s_total = merge(s_coin, s_die)
```

```{r}
head_six <- s_total %>% 
  filter(toss1 == "H", X1 == 6)

nrow(head_six) / nrow(s_total)
```

Note: this would be a different calculation than if we asked the question: what is the probability of getting a heads and then rolling a 6. We will cover that later on.

### Permutations and combinations

Permutation - an arrangement of outcomes in which the order matters

Combination - an arrangement where order does not matter

```{r}
v <- 2:5
#the below doesn't allow repeats in the combination of numbers
fac <- function(n, r) {
  factorial(n) / factorial(n - r)
}

fac(4, 4)
fac(4, 2)
fac(10, 4)
```

Imagine we are filling 3 senoir company positions(CEO, CTO, COO) from a board of 10 employees.


```{r}
library(gtools)
```

```{r}
# 2 3 4 5

permutations(4,4, repeats.allowed = TRUE)

4 * 4 * 4 * 4
```

```{r}
employees <- c("Alice", "Bob", "Carole", "David", "Elaine", "Fred", "Gabra",
               "Hua", "Inigo", "Hugo")

seniors <- as_tibble(permutations(n = 10, r = 3, v = employees,
                       repeats.allowed = FALSE))
seniors <- seniors %>% 
  rename(CEO = V1, CTO = V2, COO = V3)
```
```{r}
seniors %>% 
  count(CEO)
```



What is the probarility of Alice being selected as CEO and Bob being selected
as CTO, if the selection process is random

```{r}
seniors %>% 
  filter(CEO == "Alice", CTO == "Bob")

8/720 #8 is the number of ways that the CEO will be Alice and the CTO will be Bob and 720 is the total number of permutations of all 10 people for the 3 different positions
```

### Combinations

Permutations: you draw objects from a collectiona nd place them into a group with internal structure, e.g. a list, a hierarchy (remember our case of senior exectutives, where the three employees went into labelled positions: CEO, CTO, COO).

Combinations: you draw objects from a collection and place them into a group without internal structure, i.e. no labelled positions, roles or precedence in the group


```{r}
choose(n, r)
num_draws <- choose(50, 6)

options(scipen = 99)
1/num_draws
```

Use the combinations() function in gtools for this task (don’t try to do it 
manually). I have five flavours of ice cream to choose from:

flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio")

and a bowl that can hold three scoops. How many different variations in bowls 
of ice cream can I have assuming:

-Each scoop must be a different flavour (10)
-I can repeat flavours if I wish? (35)

```{r}
flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio")

as_tibble(combinations(5, 3, v = flavours, repeats.allowed = FALSE)) %>% 
  nrow()

as_tibble(combinations(5, 3, v = flavours, repeats.allowed = TRUE)) %>% 
  nrow()
```



## Session 3 -- Combining events

### Learning objectives

-Understand the concept of the complement of an event
-Be able to form AND and OR com

```{r}
s_four_coins <- tosscoin(4)
s_four_coins
```

What is the probability that exactly two heads are thrown? P(2 heads)

```{r}
s_four_coins <- s_four_coins %>% 
  mutate(num_heads = rowSums(. == "H"))

s_four_coins
```

```{r}
s_at_least_2_heads <- s_four_coins %>% 
  filter(num_heads >= 2)

# N 2+ heads
s_at_least_2_heads

#P 2 + heads
prob_two_heads <- nrow(s_at_least_2_heads) / nrow(s_four_coins)


prob_two_heads
1 - prob_two_heads
```

### 'And' combinations

event A: {1,2,3,4}
event B: {3,4,5,6}

What is the probability that 

```{r}
rolldie(2) %>% 
  filter(X1 %in% 3:4,
         X2 %in% 3:4)
```

### Task
Given our four coin tossing example from above, wher is the probability of tossing at least one head and at least 1 tail. To do this you'll need to add a num tails column to s_four_coins and then think about satisfying the two conditions

```{r}
s_four_coins %>% 
  mutate(num_tails = rowSums(. == "T")) %>% 
  filter(num_heads >=1, num_tails >=1) %>% 
  summarise(prob = n() / nrow(s_four_coins))
```

## Mutually exclusive events

Example:
Given the four coin tossing example,
what is the probability of tossing
    at least three heads and 
    at least two tails?

There is no way in which these two events can both happen simultaneously. We say that such events are mutually exclusive, they share no outcomes

with 4 coins
Throwing 3 heads EXCLUDES the possibility of throwing 2 tails


## OR combination

event A: roll an even number: {2, 4, 6}
event B: roll one of {4, 5}

What is the probability that A OR B occurs

```{r}
s_two_dice <- rolldie(2)
s_two_dice %>% 
  filter(X1%% 2 ==0 | X2 ==5) %>% 
  summarise(n() / nrow(s_two_dice))
```

Optional Task - 10 mins
Let’s check the addition law for ourselves. Calculate probabilities for each event below:
    first_die_even
second_die_5
first_die_even AND second_die_5
Then use these probabilities in the expression from above
prob(A∪B)=prob(A)+prob(B)−prob(A∩B)
and check you get the same probability for prob(first_die_even OR second_die_5)
as we obtained from the last block of code.

```{r}
first_die_even %>% 
  nrow()/nrow(s)

second_die_five <- s %>% 
  filter(X2 == 5)
second_die_five %>% 
  nrow()/nrow(s)

first_even_second_five <- s %>% 
  filter(X1 %% 2 == 0 & X2 == 5)

first_even_second_five %>% 
  nrow()/nrow(s)


0.5 * 0.1666667
  
0.5 + 0.1666667
```

## Session 4 -- Conditional probability

```{r}
library(janitor)
```

```{r}
cb <- read_csv("colour_blindness.csv")
```

### Learning outcomes



Probability of event A given event B

Independence

p(A|B) = p(A) --> independence - because none of the probability of A depends on event B, they are completely independent events. B can't explain any of the variance of A.


```{r}
cb %>% 
  tabyl(bio_sex, cblind) %>% 
  adorn_totals(where = c("row", "col"))

cb %>% 
  tabyl(bio_sex, cblind) %>% 
  adorn_percentages(denominator = "all") %>% 
  adorn_title()
```

```{r}
prob_female <- cb %>% 
  summarize(prob = sum(bio_sex == "female") / n())

prob_female
```

Task - 5 mins

What are the marginal probabilities of:

    being male - 60%
    being colour blind - 
    not being colour blind?
Get these probabilities from the contingency table and also calculate them using pipes

Save each in a sensibly name variable, we'll need it later

```{r}
prob_male <- cb %>% 
  summarize(prob = sum(bio_sex == "male") / n())

prob_male


cb %>% 
  tabyl(bio_sex, cblind) %>% 
  adorn_percentages(denominator = "all") %>% 
  adorn_totals(where = c("row", "col"))

prob_cb <- cb %>% 
  summarise(prob = sum(cb == TRUE) / n())
prob_cb

prob_not_cb <- cb %>% 
  summarise(prob = sum(cb == FALSE) / n())
prob_not_cb
```



```{r}
prob_female_cb <- cb %>% 
  filter(bio_sex == "female", cblind) %>% 
  summarise(prob = n()/nrow(cb))
            
prob_male_cb <- cb %>% 
  filter(bio_sex == "male", cblind) %>% 
  summarise(prob = n()/nrow(cb))
```


```{r}
prob_female * prob_cb
```

### Conditional probabilities

>What is the probability of a person being colour blind given that we know they are female?

what is the probability of throwing a 5
p(5)
What is the probability of throwing a 5, given that we know the number is 0dd
p(5|odd)



```{r}
cb %>% 
  tabyl(bio_sex, cblind) %>% 
  adorn_percentages(denominator = "all") %>% 
  adorn_totals(where = c("row", "col"))
```

p(cb) = 0.0436

p(cb | female)

> What proportion of female people are colour blind?

```{r}
prob_cb_given_female <- cb %>% 
  filter(bio_sex == "female") %>% 
  summarise(prob = sum(cblind) / n())
```

