---
title: "Day 3 - Cleaning, missing values, recoding"
output: html_notebook
---
##Session 1 -- Cleaing data
-Be introduced to the concept of Data Cleaning
-Know about some common issues that can exist within datasets
-Be able to deal with these common issues

###Packages

```{r}
library(tidyverse)
library(janitor)
```

```{r}
library(CodeClanData)
```

###Functions

'janitor::clean_names('
'rename()'

###Data

```{r}
video_games <- read_csv("steam-200k-sample.csv")
```

```{r}
video_games
```

The first 4 lines are "meta data" or data about data
```{r}
read_csv("steam-200k-sample.csv", skip = 4)
```

```{r}
video_games <- read_csv("steam-200k-sample.csv", 
         skip = 4,
         col_types = cols("bought online" = col_logical()))
video_games

read_csv("steam-200k-sample.csv", skip = 4) %>%
  mutate(`bought online` = as.logical(`bought online`))
```

```{r}
names(video_games)

video_games_janitor_clean <- video_games %>% 
  clean_names()
video_games_janitor_clean
```

```{r}
video_games_janitor_clean %>% 
  rename("game_title" = "game_name", 
         "activity" = "what",
         "playing_hours" = "value")
```

##Session 2 -- Missing values

- Understand the challenges of missing and null values
    - missing is not the same as null
- Know potential methods of dealing with missing values
- Identify why the data was missing - random or pattern
- Be able to identify and justify the approach taqken for each missing data item
- Be able to tidy up a dataset and deal with missing values

### Functions
is.na()
drop_na()
coalesce() <--> na_if() #They are complements of each other
na_if()
as.numeric()
as.****()

Sneak peak at "ggplot2"

### Data

```{r}
comms_data
```

MCAR - Missing completely at random - analyses can accomodate
MAR - Missing at random - maybe okay, but the missing data could be related to another confounding variable
MNAR - Bad news, the data is missing due to an underlying pattern within the same variable
NaN - Not a number - e.g. 0/0
Inf - Infinity
NULL - complete lack of a value
"Unknown" - character representation of an unknown value
NA - How we explicitly encode missingness


###Drop all NAs
but this could lead to losing a lot of data
```{r}
comms_data %>% 
  drop_na()
```

#Imputation
Replace the NAs with ... something sensible
Maybe the mean or the median
```{r}
comms_data %>% 
  pull(monthly_charges) %>% 
  mean(na.rm = TRUE)

comms_data %>% 
  pull(monthly_charges) %>% 
  median(na.rm = TRUE)

min_charge <- comms_data %>% 
  pull(monthly_charges) %>% 
  min(na.rm = TRUE)

max_charge <- comms_data %>% 
  pull(monthly_charges) %>% 
  max(na.rm = TRUE)

max_charge - min_charge

summary(comms_data)
```

## Leave them alone?
Most functions in R have a default way of dealing with NAs, so we need to be aware
of how they are handled by default

"is.na()"

```{r}
comms_data
```

##Standard missing values

'NA'


```{r}
comms_data
comms_data %>% 
  filter(is.na(monthly_charges))

#Tells you the percentage of missing data
comms_data %>% pull(monthly_charges) %>% 
  is.na() %>% 
  mean()


comms_data %>% 
  summarise(count = sum(is.na(monthly_charges)))
```

##Imputation
coalesce(thing to fix, thing to fix with)

```{r}
comms_data_imputed_median <- comms_data %>% 
  mutate(monthly_charges)

comms_data %>% 
  select(monthly_charges) %>% 
  mutate(monthly_charges2 = coalesce(monthly_charges, 
                                     median(monthly_charges, na.rm = TRUE)))

summary(comms_data)
```

```{r}
comms_data_imputed_values <- comms_data %>% 
  mutate(monthly_charges_median = coalesce(monthly_charges, 
                                           median(monthly_charges, na.rm = TRUE)),
         monthly_charges_mean = coalesce(monthly_charges, 
                                         mean(monthly_charges, na.rm = TRUE)),
         .after = monthly_charges)

comms_data_imputed_values %>% summary
comms_data %>% summary
```

```{r}

ggplot(comms_data_imputed_values, aes(customerID, monthly_charges_median)) +
  geom_point()

ggplot(comms_data_imputed_values, aes(customerID, monthly_charges_mean)) +
  geom_point()
```

### Non-standard missing values

Take a look

```{r}
comms_data %>% 
  pull(payment_method)
```

```{r}
comms_data %>% 
  pull(total_charges)
```

'na_if() -- kind of like coalesce, but instead of taking NAs and converting them into a value, it
converts different strings into NA 

```{r}
comms_data_imputed_values %>% 
  select(total_charges) %>% ##can delete safely, just for demonstration
  mutate(total_charges_explicit_na = na_if(total_charges, "na"),
         total_charges_explicit_na2 = na_if(total_charges, "N/A"))
```

```{r}
comms_data_imputed_values %>% 
  mutate(total_charges = na_if(total_charges, "na"),
         total_charges = na_if(total_charges, "N/A"))
```

```{r}
comms_data_cleaned <- comms_data %>% 
  mutate(total_charges = na_if(total_charges, "na"),
         total_charges = na_if(total_charges, "N/A"),
         payment_method = na_if(payment_method, "--"),
         payment_method = coalesce(payment_method, "unknown"))
comms_data_cleaned
```

##Recoding

-Understanding _what_ recoding is
-Know how to use the functions `if_else`, `recode` and `case_when`
-Understand _when_ to use each function.

###Data

```{r}
CodeClanData::students_big
```

### `if_else(`)
-`if_else(TEST, value if TRUE, value if FALSE)`

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(older_than_12 = ageyears > 12)
```
```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(older_than_12 = if_else(ageyears > 12, "Older than 12", "12 or younger"))
```


```{r}
students_big %>% 
  select(gender) %>% 
  mutate(gender_recoded = if_else(gender == "M", "Male", "Female"))
```

```{r}
students_big %>% 
  mutate(handed_recoded = if_else(handed == "R", "Right", "Left"),
         favourite_physical_activity_recoded = if_else 
         (favourite_physical_activity == "Football/Soccer", "Football", favourite_physical_activity))
```

##`recode()`

```{r}
students_big %>% 
  select(travel_to_school) %>% 
  mutate(travel_to_school_recoded = recode(travel_to_school,
                                           "Bus" = "bus",
                                           "Car" = "car",
                                           "Walk" = "walk"))
students_big %>% 
  count(travel_to_school)

students_big %>% 
  select(travel_to_school) %>% 
  mutate(travel_to_school_recoded = recode(travel_to_school,
                                           "Bus" = "bus",
                                           "Car" = "car",
                                           "Walk" = "walk",
                                           .default = "other"))
```

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(ageyears_recoded = recode(ageyears, "12" = 120,
                           .default = 0))
```


```{r}
data <- read.table(header = TRUE, text="
 subject sex control cond1 cond2 compl
       1   M     7.9  12.3  10.7  T
       2   F     6.3  10.6  11.1  T
       3   F     9.5  13.1  13.8  N
       4   M    11.5  13.4  12.9  F
")
```

```{r}
data %>% 
  mutate(subject = recode(subject,
                          "1" = "001",
                          "2" = "002",
                          "3" = "003",
                          "4" = "004"),
         sex = recode(sex, 
                      "M" = "Male",
                      "F" = "Female"),
         compl = recode(compl, "T" = "True",
                        "F" = "False",
                        .default = NA_character_))
```

##`case_when()`

```{r}
x <- 1:10

case_when(x <= 5 ~ "A",
          x<= 7 ~ "B",
          TRUE ~ "C")
```

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(
    age_divisible = case_when(
      ageyears %% 2 == 0 ~ "Even",
      ageyears %% 3 == 0 ~ "Divisible by 3",
      ageyears %% 5 == 0 ~ "Divisible by 5",
      TRUE               ~ "Other number"
    )
  )
```

