---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(car)
library(modelr)
library(skimr)
library(GGally)
library(ggfortify)

skim(Prestige)
```

# Manual models

```{r}
 prestige_trim <- Prestige %>% 
 drop_na() %>% 
  select(-census)
```

## Feature engineering: logged some positively skewed numerics

```{r}
prestige_trim <- prestige_trim %>% 
  mutate(ln_women = log(1 + women),
         ln_income = log(income))
```

```{r}
prestige_trim %>% 
  select(-c(ln_women, ln_income)) %>% 
  select(prestige, everything()) %>% 
  ggpairs(aes(colour = type, alpha = 0.5))
```

```{r}
prestige_trim %>% 
  select(prestige,ln_women, ln_income, type) %>% 
  #select(prestige, everything()) %>% 
  ggpairs(aes(colour = type, alpha = 0.5))
```

```{r}
prestige_trim %>% 
  slice_max(income, n = 10)
```

```{r}
mod1a <- lm(prestige ~ education,
            data = prestige_trim)

summary(mod1a)
```

```{r}
autoplot(mod1a)
```

```{r}
mod1b <- lm(prestige ~ type, data = prestige_trim)

autoplot(mod1b)

summary(mod1b)
```

type bc has an average prestige value of 35.52

For professional type jobs, we can expect to see, on average, an increase of 32.32 in prestige score compared to blue collar type jobs
For white coller type jobs, we can expect to see, on average, an increase of 6.7 in prestige score compared to blue collar type jobs
The model explains 70% of the variation in prestige score


Now we want to see which variables describe the "residue", the unexplained model error

```{r}
prestige_resid <- prestige_trim %>% 
  add_residuals(mod1a) %>% 
  select(-c(prestige, education))

prestige_resid %>% 
  select(resid, everything()) %>% 
  ggpairs(aes(colour = type, alpha = 0.5))
```

## add a second predictor: next best 
i.e. the one that explains the most residual error

```{r}
mod2a <- lm(prestige ~ education + ln_income, 
            data = prestige_trim)

summary(mod2a)

autoplot(mod2a)
```

```{r}
mod2b <- lm(prestige ~ education + income,
            data = prestige_trim)

summary(mod2b)
```

```{r}
mod2c <- lm(prestige ~ education + type,
            data = prestige_trim)

summary(mod2c)
```

mod2a gave the biggest uplift in r2 and residuals look great - add it in to the model


## check of significance of categorical: ANOVA test
```{r}
anova(mod1a, mod2c)
```

it is statistically significant (i.e. the 3 levels have different means), so we can include type in the model with education


## third predictor!
```{r}
prestige_resid <- prestige_trim %>% 
  add_residuals(mod2a) %>% 
  select(-c(prestige, education, ln_income))

prestige_resid %>% 
  select(resid, everything()) %>% 
  ggpairs(aes(colour = type, alpha = 0.5))
```

```{r}
mod3a <- lm(prestige ~ education + ln_income + women,
            data = prestige_trim)

summary(mod3a)
```

```{r}
mod3b <- lm(prestige ~ education + ln_income + type,
            data = prestige_trim)

summary(mod3b)
```

```{r}
autoplot(mod3b)
```

```{r}
anova(mod2a, mod3b)
```

Diagnostic plots are fine, anova test is significant, let's add type to the model.

## adding an interaction (between existing main effects)

```{r}
prestige_resid <- prestige_trim %>% 
  add_residuals(mod3b) %>% 
  select(-prestige)
```

```{r}
coplot(resid ~ ln_income | education,
       panel = function(x, y, ...){
         points(x, y)
         abline(lm(y ~ x), col = "blue")
       },
       data = prestige_resid,
       rows = 1)
```

The effects of income are different for different levels of education, so it could be worth adding them into the model

```{r}
prestige_resid %>% 
  ggplot(aes(x = education,
             y = resid,
             colour = type))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```

The effects of education are different for different types of workers 

```{r}
prestige_resid %>% 
  ggplot(aes(x = ln_income,
             y = resid,
             colour = type))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
mod4a <- lm(prestige ~ education + ln_income + type +
              education:ln_income,
            data = prestige_trim)

summary(mod4a)
```

```{r}
mod4b <- lm(prestige ~ education + ln_income + type +
              education:type,
            data = prestige_trim)

summary(mod4b)
```

```{r}
mod4c <- lm(prestige ~ education + ln_income + type +
              ln_income:type,
            data = prestige_trim)

summary(mod4c)
```

```{r}
autoplot(mod4c)
```

```{r}
anova(mod3b, mod4c)
```

```{r}
#this breaks down the relative importance of each of the predictors - it breaks down the r^2 value 
library(relaimpo)

calc.relimp(mod4c, type = "lmg", rela = TRUE)
```

