---
title: "R Notebook"
output: html_notebook
---

# Joining data lesson

## Table Relationships

-One-to-one - for each row in the primary table, there will be a single row in 
the foreign table
-One-to-many - Rows from the foreign table can correspond to multiple rows
in the primary key, but each row in primary key only corresponds to one row
in the foreign key
-Many-to-many - multiple rows can correspond to multiple rows, but must be 
tracked through an intermediary table b/c primary key must be unique

## Keys

-Primary key columns - those which establish a unique identifier for a table
-Foreign key columns - columns in a table that contain values corresponding to 
the primary keys of another table
  -Such as using numbers to represent different categories (e.g. 1 for high
  school, 2 for some college, 3 for university degree, 4 for postgraduate
  degree)
  
# library
```{r}
library(tidyverse)
```

# data
```{r}
customers <- tibble(
  customer_ID = c(001, 002, 004, 005, 008, 010), 
  name = c("John Smith", "Jane Adams", "Robert Landry", "Jane Tow", "Raul Tann", "Hilary Joyal"),
  email_address = c("johnsmith@gmail.com", "janea@gmail.com", "rlabdry@hotmail.com", "janet89@aol.com", "paul.tann1@gmail.com", NA),
  shipping_address = c("32 Station Road, Edinburgh", "42 Park Drive, Edinburgh", NA, "10 Gardiner Rd, Edinburgh", "12 Main St, Edinburgh", " 234 Queensferry Rd, Edinburgh,")
)

customers
```

```{r}
orders <- tibble(
  customer_ID = c(001, 002, 005, 010, 012, 001), 
  order_number = c("KU65781", "SU98201", "JS76821", "YR28212", "JS91221","YR21873"),
  order_total = c("£32.50", "£56.00", "£100.02", "£20.25", "£39.50", "£45.00"),
  product_code = c("XZY780", "XYZ123", "XYZ256", "XYZ262", "XYZ345", "XYZ127")
)

orders 
```

# Join datasets

4 mutating join verbs:
* inner_join() : returns all rows from two tables with matching records.
* left_join() : returns all records in left table, and any matching records in another.
* right_join() : returns all records in right table, and any matching records in another.
* full_join() : returns all records in both tables, matching where possible.

2 filtering join verbs:
* semi_join() : returns only rows from the first table where one or more matches are found in the second table.
* anti_join() : returns only rows from the first table where no matches are found in the second table

2 binding verbs:
* bind_rows() : add rows to a table
* bind_cols() : add columns to a table

```{r}
#join(datax, datay, c(keyx, keyy))
```

```{r}
inner_join(customers, orders, "customer_ID")
```

```{r}
left_join(customers, orders, "customer_ID")
```

```{r}
right_join(customers, orders, "customer_ID")
```

```{r}
full_join(customers, orders, "customer_ID")
```

```{r}
customers_adj <- rename(customers, ID = customer_ID)
```

```{r}
inner_join(customers_adj, orders, by = c("ID" = "customer_ID"))
```

```{r}
customers2 <- customers %>% 
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "card"))

orders2 <- orders %>% 
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "card"))
```

```{r}
inner_join(customers2, orders2, by = c("customer_ID", "payment_type"))
```

```{r}
semi_join(customers, orders, "customer_ID")
```

```{r}
anti_join(customers, orders, "customer_ID")
```

```{r}
# SLACK THIS OUT
statuses <- tibble(
  status = c("active", "active", "inactive", "inactive", "active", "inactive")
) 
```

```{r}
#When using bind_cols, both datasets must be in the same order.
bind_cols(customers, statuses)
```

```{r}
past_customers <- tibble(
  customer_ID = c(020, 021), 
  name = c("Dave Brown", "Jess Bishop"),
  email_address = c("dbrown@hotmail.com", "jessb@gmail.com"),
  address = c("57 Park View, Leith", "12 Reid Way, Broxburn")
)

past_customers
```

```{r}
#Doesn't need all the same column names
bind_rows(customers, past_customers)

#Needs all the same column names, so will return error if column names aren't
#exactly the same in both tables.
rbind(customers, past_customers)
```

# Joins lab

# Data
```{r}
movies <- read_csv("movies.csv")
roles <- read_csv("roles.csv")
actors <- read_csv("actors.csv")
```
#1.1
## Question 1
What are the primary and foreign keys in each data frame
Actors dataset - 
  -Primary key: ID
  
Movies dataset - 
  -Primary key: id

Roles dataset - 
  - Foreign keys - movie ID, actor ID


## Question 2
- Relationship btwn actors and roles
Actor ID in roles corresponds to primary key in actors dataset - one to many
-movies and roles
Movie ID in roles corresponds to primary key in movies dataset - one to many
- actors and movies
Many to many

# 1.2
## Question 1: From the movies table, find the id of the film The Little Mermaid. (In the data as “Little Mermaid, The”).
```{r}
movies %>% 
  filter(name == "Little Mermaid, The")
```
## Question 2: Filter the roles data frame, so we just have roles from the Little Mermaid (hint: use the ID you found in the last question). Save the result as little_mermaid_roles.


```{r}
little_mermaid_roles <- roles %>% 
  filter(movie_id == "192017")
```
## Question 3: Carry out a left join on this new data frame and the actors data frame, to find the actors which appeared in The Little Mermaid.

```{r}
left_join(little_mermaid_roles, actors, by = c("actor_id" = "id"))
inner_join(little_mermaid_roles, actors, by = c("actor_id" = "id"))
```
# 1.3
## Question 1: For all actors who have appeared in at least two films. Call this data frame actors_more_than_once.
```{r}
actors_more_than_once <- actors %>% 
  filter(film_count >= 2)
```
## Question 2: Left join actors_more_than_once to roles, so we can see all the actors roles those actors played in.

```{r}
left_join(actors_more_than_once, roles, by = c("id" = "actor_id"))
```

## Question 3: Left join the result from above to movies, so we can see details of each movie for these roles
```{r}
actors_more_than_once %>% 
left_join(roles, by = c("id" = "actor_id")) %>% 
  left_join(movies, by = c("movie_id" = "id"))
```


## Question 4: For each actor, find the average rating of the films they appeared. Find the five actors with the highest average rank of films they appeared in.

```{r}
actors_more_than_once %>% 
left_join(roles, by = c("id" = "actor_id")) %>% 
  left_join(movies, by = c("movie_id" = "id")) %>% 
  #changes rank into a numeric variable
  mutate(rank = as.numeric(rank)) %>% 
  #groups by both first name and last name
  group_by(first_name, last_name) %>% 
  #summarise mean for each actor
  summarise(avg_rank = mean(rank, na.rm = TRUE)) %>%
  #ungroups
  ungroup() %>% 
  #Takes the top 5
  slice_max(avg_rank, n = 5)
```

#Tidy data lesson
## Functions
  -pivot_longer() - gathers columns into rows
  -pivot_wider(): speads rows into columns
  -separate(): separates one column into two
  -unite(): unites several columns into one
  
##Examples

```{r}
subject_hr <- tibble( name = c("SUBJ01","SUBJ02","SUBJ03"), 
  hr_am = c(68,72,68), 
  hr_pm = c(58,89,52))

subject_hr
```
This data is in wide format - each variable doesn't have its own column

```{r}
subject_hr %>% 
  pivot_longer(cols = c("hr_am", "hr_pm"), 
               names_to = "measurements_of_time", 
               values_to = "heartrate")
```

```{r}
income <- tibble(religion = c("Agnostic","Atheist","Buddhist", "Catholic"), 
  X10k = c(12,25,3,75), 
  X10k_to_20k = c(34,14,6,60),
  X20k_to_30k = c(45,22,10,110))

income
```
```{r}
income %>% 
  pivot_longer(cols = starts_with("X"),
               names_to = "income_bracket",
               values_to = "income")
```
```{r}
messy_orders <- tibble( name = c("Joe Bloggs","Jane Smith","Penny Jones"), 
  order_info = c("XY7282/29-06-19","LO0182/15-03-19","AS6812/04-06-19"), 
  order_amount = c(58.00,125.00,75.25))

messy_orders
```
```{r}
tidy_orders <- messy_orders %>% 
  separate(order_info, 
           into = c("order_number", "date"),
           sep = "/")
```

```{r}
library(CodeClanData)
```
```{r}
hospital_visits_long <- hospital_visits %>% 
  pivot_longer(cols = starts_with("FY"), 
               names_to = "year", 
               values_to = "visit_number",
               #removes names' prefixes
               names_prefix = "FY")
```

```{r}
tidy_hospital <- hospital_visits_long %>% 
  pivot_wider(names_from = Field, 
              values_from = visit_number) %>% 
  separate(IcdChapter, c("code", "description"), sep = "\\.")
```

```{r}
tidy_hospital %>% 
  unite(disease, c("code", "description"), sep = ":")
```

# Tidy data lab
```{r}
country_data <- read_csv("country_data.csv")
```

What needs to change to make it tidy
-population and cases should be in separate columns
-Year should be in its own column
-The numbers should be under the columns population and cases

hospital_visits_long <- hospital_visits %>% 
  pivot_longer(cols = starts_with("FY"), 
               names_to = "year", 
               values_to = "visit_number",
               #removes names' prefixes
               names_prefix = "FY")

```{r}
country_data_long <- country_data %>% 
  pivot_longer(cols = starts_with("Y"),
               names_to = "year",
               values_to = "pop_count",
               names_prefix = "Y")
```

tidy_hospital <- hospital_visits_long %>% 
  pivot_wider(names_from = Field, 
              values_from = visit_number)
```{r}
tidy_country <- country_data_long %>% 
  pivot_wider(names_from = type,
              values_from = pop_count) %>% 
  separate(col = country, 
           c("country_number", "country_name"), 
           sep = "\\.") %>% 
  unite(country_code, c(code_letter, code_numb), sep = "")
```

