---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```
##Part 1 - Coding best practices
```{r}
amount_spent <- c(9.1, 8.55, 14.32, 14.17, 14.34, 7.97, 9.79, 11.47, 12.44)
category <- c(
  "lunch",
  "dinner",
  "lunch",
  "lunch",
  "cinema",
  "drinks",
  "lunch",
  "lunch",
  "dinner"
)
my_week_of_spending <- tibble(
  amount_spent      = amount_spent,
  spending_category = category
)
summarise(my_week_of_spending, sum(amount_spent))
summarise(
  filter(
    my_week_of_spending, spending_category == "lunch"
  ),
  sum(amount_spent)
)
```

##Part 2 - Errors

#checking type of data
```{r}
class(mtcars)
```
#Conflicting functions
Such as using #library(dplyr) and #library(MASS) at the same time


#B/c both MASS and dplyr have a function called select, you have to specify 
which you want to run since one is masking the other
```{r}
mtcars %>% 
  dplyr::select(mpg)
```
##Show conflicts
```{r}
conflicts()
```

##Errors lab

##Error 1
#Original error
some_letters <- ("E", "R", "R", "O", "R)

#fixed
```{r}
some_letters <- c("E", "R", "R", "O", "R")
some_letters
```

##Error 2
#Original error
message <- c("does", "this" "work", "?")

toupper(messege)

#fixed
```{r}
message <- c("does", "this", "work", "?")

toupper(message)
```

##Error 3
#Original error
mtcars %>% 
  select(mpg, cyl, disp) %>>%
  filter(mpg < 20)

#Fixed
```{r}
mtcars %>% 
  select(mpg, cyl, disp) %>%
  filter(mpg < 20)
```

##Error 4
#Original error
iris %>% 
  select(Sepal.Width, Species) %>% 
  arrange(Sepal.Length)
  
#Fixed
```{r}
iris %>% 
  select(Sepal.Length, Sepal.Width, Species) %>% 
  arrange(Sepal.Length)
```

##Error 5
#Original error
ToothGrowth %>%
  summarise(avg_len_vc = mean(len)) %>% 
  filter(supp == "VC")
  
#Fixed
```{r}
ToothGrowth %>%
  filter(supp == "VC") %>% 
  summarise(avg_len_vc = mean(len))
```



##Error 6
#Original error
animals <- c("cat", dog, "rabbit", "horse")
sample(animals, 10)

#Fixed
```{r}
animals <- c("cat", "dog", "rabbit", "horse")
sample(animals, 10, replace = TRUE)
```

##Error 7
#Original error
dog_names <- c("Lucy", "Molly", "Charlie", "Luna", "Bella", "Cooper")

cat_names < c("Luna", "Lily", "Lucy", "Charlie", "Max", "Oliver")
Intersect(dog_names)

#Fixed
```{r}
dog_names <- c("Lucy", "Molly", "Charlie", "Luna", "Bella", "Cooper")

cat_names <- c("Luna", "Lily", "Lucy", "Charlie", "Max", "Oliver")

intersect(dog_names, cat_names)
```

##Error 8 
#Original error
chickwts %>% 
  filter(weight =< 150) %>% 
  mutate(weight_kg == weight * 0.001)
  
#Fixed
```{r}
chickwts %>% 
  filter(weight <= 150) %>% 
  mutate(weight_kg = weight * 0.001)
```

##Error 9
#Original error
tibble(alphabet = LETTERS(),
       numbers = 1:25) %>% 
  filter(letter != c(A, B, C))
  
#Fixed
```{r}
tibble(alphabet = LETTERS,
       numbers = 1:26) %>% 
  filter(alphabet != c("A", "B", "C"))
```

##Error 10
```{r}
set.seed(999)

scent_ratings <- tibble(
  iD = rep(1:10, each = 5),
  SCENT = rep(c("New book", "Cut grass", "Rose oil", "Fresh linen", "Wet dog"), 10),
  rating = sample(c(0:5, NA), 50, replace = TRUE),
  PersonAge = rep(NA, 50)
)
```

```{r}
scent_ratings <- janitor::clean_names(scent_ratings)

```
```{r}
scent_ratings %>% 
  drop_na(rating) %>% 
  group_by(scent) %>% 
  summarise(mean_rating = mean(rating))
```

