---
title: "R Notebook"
output: html_notebook
---

# Libraries

```{r}
library(tidyverse)
library(janitor)
```
# Question 1. Read in the data containing dietary compositions and familiarise yourself with it.

```{r}
country_codes <- read_csv("country_codes.csv")
dietary_by_country <- read_csv("dietary-composition-by-country.csv")
country_codes
dietary_by_country
```

# Question 2 - Change the data to long format with food categories going to a column called kcal_source and the calorie values going to a column called avg_daily_kcals. Save into variable diet_comp_clean

```{r}
diet_comp_clean <- dietary_by_country %>% 
pivot_longer(cols = ends_with("(FAO (2017))"),
names_to = "kcal_source",
values_to = "avg_daily_kcals")
diet_comp_clean
```

# Question 3 - Clean kcal_source categories by removing any unnecessary information. Then clean all column names, and rename the column ‘entity’ to ‘country’. Overwrite diet_comp_clean with your results. [Hint: you’ll probably have to use some regex to clean kcal_source categories]

```{r}
diet_comp_clean %>%
  mutate(kcal_source = str_sub(kcal_source, 1, nchar(kcal_source) - 13)) %>% 
  clean_names() %>% 
  rename(country = entity)

```

